A Study of Common Bug Fix Patterns in Rust \thanks{Competing Interests: We have no competing interests and are funded by a Discovery Grant from Canada's Natural Science and Engineering Research Council.}

### 2023-06-14 conclusion.tex
capitalization
last read through

In this project, we developed Ruxanne, a tool for mining bug fix patterns in Rust. Using this tool, we mined the top 18 most starred Rust projects in Github to discover common bug fix patterns within their code changes. At the heart of Ruxanne, we used a novel code embedding approach to embed the most important aspects of a code change in a fixed sized datapoint. We stored 87,726 datapoints drawn from 57,214 commits across these 18 projects. After using the DBSCAN clustering algorithm, and a subsequent manual analysis, we obtained 12 clusters of general fix patterns and 8 clusters of borrow-checker related fix patterns.

>>>>

In this project, we developed Ruxanne, a tool for mining bug fix patterns in Rust. Using this tool, we mined the top 18 most starred Rust projects in GitHub to discover common bug fix patterns within their code changes. At the heart of Ruxanne, we used a novel code embedding approach to embed the most important aspects of a code change in a fixed sized datapoint. We processed 87,726 datapoints drawn from 57,214 commits across these 18 projects. After using the DBSCAN clustering algorithm, and a subsequent manual analysis, we obtained 12 clusters of general fix patterns and 8 clusters of borrow-checker related fix patterns. We found that the most common general pattern was adding or removing fields, while the most common borrow-checker related pattern was removing a \texttt{clone()} call.

### 2023-06-14 discussion.tex
reedit
last read through

When presenting each pattern, we discussed actionability for the most important actions. Here, we discuss the implications of our most important patterns. \emph{In summary, program repair tools must understand semantics of attributes and should be performance-aware.} It may also be useful for tools to be able to automatically infer immutability. Here, we discuss the top ten most popular patterns, 8 of which fall into the category of general patterns, and the remaining 2 are BC-related patterns.

@@

The most common borrow-checker-related pattern and second-most-common pattern overall is dropping \texttt{clone()}. This pattern is performance-related; when correctly applied, it has no semantic impact on the program, but reduces resource consumption. Rust is used in situations where performance is important; the number of instances of this pattern confirms that developers do work on performance issues. This implies that \emph{program repair tools would benefit from being performance-aware}, rather than limiting themselves strictly to functional properties of the code.

@@

\vspace*{1em}
\fbox{\parbox{0.9\textwidth}{We summarize learnings from our bug patterns as follows:
\begin{itemize}
\item Rust program repair tools must understand the effects of attributes.
\item Program repair tools should be performance-aware.
\item Tools for inferring Rust immutability could be viable.
\end{itemize}
}}

@@

A threat to external validity is selection bias, where the benchmarks differ systematically from the actual world. The selected projects might cause selection bias and therefore we might not have presented patterns that may exist in other projects. Another threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing a functionality. Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes.

>>>>

When presenting each pattern, we discussed actionability for the most important actions. Here, we discuss the implications of our most important patterns. \emph{In summary, program repair tools must understand semantics of attributes and should be performance-aware.} It may also be useful for tools to be able to automatically infer immutability. Here, we discuss the top ten most popular patterns (8 general patterns and 2 BC-related patterns).

@@

The most common borrow-checker-related pattern and second-most-common pattern overall is dropping \texttt{clone()}. This pattern is performance-related; when correctly applied, it has no semantic impact on the program, but reduces resource consumption. Rust is used in situations where performance is important, and the number of instances of this pattern confirms that developers do work on performance issues. This implies that \emph{program repair tools would benefit from being performance-aware}, rather than limiting themselves strictly to functional properties of the code.

@@

\vspace*{1em}
\fbox{\parbox{0.9\textwidth}{We summarize some learnings from our bug patterns:
\begin{itemize}
\item Rust program repair tools must understand the effects of attributes.
\item Program repair tools should be performance-aware.
\item Tools for inferring Rust immutability could be viable.
\end{itemize}
}}

@@

A threat to external validity is selection bias, where the benchmarks differ systematically from the actual world. Our selection of projects might be biased and therefore we might not have presented patterns that may exist in other projects. Another threat to external validity can be the developers reporting the commits as bug fixing commits while they are not really fixing functionality. Similarly, a commit message might not contain our target keywords while the commit is associated with bug fixing changes.

### 2023-06-14 methodology.tex
last read through

We define a code change to be a modification to a program's source code that changes the program's abstract syntax tree (AST). A change pattern is a set of code changes that serves a specific purpose. Purposes include rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns) and changes that support bug-fixing patterns (fix-induced patterns) as well as improving code readability or maintainability (refactoring patterns). In this work, we are interested in bug-fixing and fix-induced patterns. Thus, from now on, whenever we refer to code patterns, we mean bug-fixing and fix-induced patterns.

Our approach for automatically finding pervasive code patterns aims to find clusters of similar code patterns, using existing implementations of appropriate clustering algorithms. Work in this vein generally makes the assumption that clusters corresponds to classes of change patterns~\citep{hanam2016discovering,campos2019discovering,yang2022mining}. We thus want to associate code changes with datapoints; one of our contributions is a fixed-size embedding which highlights the important parts of a code change.

@@

Syn\footnote{\url{https://crates.io/crates/syn}} is a Rust crate built for the Rust procedural macro implementation. However, it includes a parser which is suitable for our purposes; we simply had to write a preprocessor to transform Syn AST output into Python dictionaries. Syn handles all of Rust. Each of our datapoints specifies whether any Syn non terminal (\texttt{NT}) is present or absent, i.e. containing one dimension per non terminal. However, we are also interested in change patterns that involve the borrow checker (BC), so we also add a dimension reflecting the presence of a set of BC-related elements (\texttt{BCE}) that we identified, e.g. \texttt{clone}, \texttt{Rc}, \texttt{Box}. The full list of \texttt{BCE} elements can be found in our replication package\footnote{\url{https://zenodo.org/record/7388618}}.

@@

The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occured. The third part is the content of the change---terminals and nonterminals. Ruxanne looks for changes within each root-level scope (the Rust book\footnote{\url{https://doc.rust-lang.org/reference/items.html}} calls these root-level scopes the \textit{items} of a crate). That is, our assumption is that each change pattern occurs within one root-level scope. In this work, we chose to not handle patterns that involve changes in multiple functions or multiple files. However, our tool detects both continuous and non-continuous line changes within one scope.

@@

\paragraph{Example: path extraction.} Figure~\ref{fig:extraction} illustrates path extraction on a simple change, starting from the ASTDiff. The change (shown in the top right corner of the Figure) replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+. After the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, the third argument sends a clone of \verb+head+ to the callee instead of sending \verb+head+ itself. A DFS on this tree yields three different paths, which we show in red, green, and purple. A path represents a sequence of involved non terminals. It is crucial to record the order of nodes in the path, so that we can record that order in our datapoint embedding.

@@

The most naive approach would be to only report the observed non terminals within the diff. That yields a fully order-insensitive representation. In such an embedding, for instance, there is no difference between two nested if statements and two if statements beside each other. On the other hand, theoretically, for full order-sensitiveness, we would need to embed all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields a combinatorial explosion; this could potentially be reduced somewhat, but is still impractical. While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.

A reasonable work around would be to pass from the full set of non terminals to a smaller set of categories (e.g. a category for larger entites like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to \cite{hanam2016discovering}). That ad-hoc approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations. However, it would still yield a sparse dataset.

We propose a novel representation. Similar to the naive approach, we define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ collected from Syn. However, to account for order-sensitivity, we introduce two additional steps. Firstly, we record the number of occurrences of items from $\mathtt{NT} \cup \mathtt{BCE}$ within each distinct root-level scope (e.g. structs, functions, impl blocks, etc). This captures the varying occurrence counts resulting from different element orders in the program. Secondly, we incorporate a weighting scheme, designed semi-automatically, to prioritize non-terminals that we consider more salient for Rust bugs.

\paragraph{Example: computing fixed-sized vectors.} Figure~\ref{fig:essence} depicts the fixed-sized vectors, as described in this subsection, for the change from Figure~\ref{fig:extraction}. All of the columns in this vector belong to scope \verb+ItemFn+---that is, these non-terminals occur within a function. The Figure also depicts the essence of change, to be described immediately below.

@@

    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights results in the essence of change, visualized with a circle pack figure at bottom right. All non terminals here occur within scope {\texttt ItemFn}.}
\end{figure}

Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff. For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments. The last description, in our opinion, is the most useful one, and we designed our weighting scheme to embed this value judgment.
We made two observations about changes. First, we saw that items (non terminals or BC-related elements) that occurs closer to AST leaves tend to be more semantically important. Second, and relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g. ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}). We thus want to prioritize items that tend to occur closer to leaves and de-prioritize items that tend to occur closer to roots.

To design our weighting scheme, we collected empirical data about locations of item occurrences in a corpus. Specifically, we mined 20 of the most recent bug related commits of the same projects described in Table~\ref{table:repos}, and ran them through our pipeline. We recorded the total number of occurrences of each item. Then, if $\# i$ is the number of occurences of $i$, we gave $i$ a weight of $1/\# i$. Because items that occur closer to the root also occur more often, per our observation, this gives higher weights to items that occur infrequently, i.e. closer to the leaves. Items that occur often and have lower weights are, we believe, less important in describing a change.

@@

\paragraph{Example: essence of change.} Figure~\ref{fig:essence} shows a complete example for computing the essence of change. The essence of change vector essentially serves as our embedding vector. The Figure shows the process for transforming the patch illustrated in Figure~\ref{fig:extraction} from fixed-size datapoints (computed as described in the previous subsection) into the essence of change vector. To better illustrate the essence vector, Figure~\ref{fig:essence} also graphically presents it using a circle pack figure. The radius of each circle corresponds to the importance of the item inside it.

@@

\subsubsection{\label{sec:manual_analysis_parameter_tuning}Manual Analysis for parameter tuning}

@@

\subsubsection{\label{sec:manual_analysis_cluster_selection}Manual Analysis for cluster selection}
% 3 types, refactoring dropped
Like \cite{yang2022mining}, and following the definitions of \cite{cotroneo2019analyzing}, we divide the clusters in three different groups:

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
// First Pattern
-- foo1(arg.clone);
++ foo1(&arg);

// Second Pattern
fn foo2() {
--  stmt1;
++  stmt2;
}

>>>>

We define a code change to be a modification to a program's source code that changes the program's abstract syntax tree (AST). A change pattern is a set of code changes that serves a specific purpose. Purposes include rectifying program behaviour with respect to a specific functional or non-functional requirement (bug-fixing patterns) and changes that support bug-fixes (fix-induced patterns) as well as improving code readability or maintainability (refactoring patterns). In this work, we are interested in bug-fixing and fix-induced patterns. Thus, from now on, whenever we refer to code patterns, we mean bug-fixing and fix-induced patterns.

Our approach for automatically finding pervasive code patterns aims to find clusters of similar code patterns, using existing implementations of appropriate clustering algorithms. Work in this vein generally makes the assumption that clusters correspond to classes of change patterns~\citep{hanam2016discovering,campos2019discovering,yang2022mining}. We thus want to associate code changes with datapoints; one of our contributions is a fixed-size embedding which highlights the important parts of a code change.

@@

Syn\footnote{\url{https://crates.io/crates/syn}} is a Rust crate built for the Rust procedural macro implementation. However, it includes a parser which is suitable for our purposes; we simply had to write a preprocessor to transform Syn AST output into Python dictionaries. Syn handles all of Rust. Each of our datapoints specifies whether any Syn non~terminal (\texttt{NT}) is present or absent, i.e. containing one dimension per non~terminal. However, we are also interested in change patterns that involve the borrow checker (BC), so we add a dimension reflecting the presence of a set of BC-related elements (\texttt{BCE}) that we identified, e.g. \texttt{clone}, \texttt{Rc}, \texttt{Box}. The full list of \texttt{BCE} elements can be found in our replication package\footnote{\url{https://zenodo.org/record/7388618}}.

@@

The second part identifies the context of the modification, that is, the path from the root of the tree to the subtree in which the modification has occurred. The third part is the content of the change---terminals and non~terminals. Ruxanne looks for changes within each root-level scope (the Rust book\footnote{\url{https://doc.rust-lang.org/reference/items.html}} calls these root-level scopes the \textit{items} of a crate). That is, our assumption is that each change pattern occurs within one root-level scope. In this work, we chose to not handle patterns that involve changes in multiple functions or multiple files. However, our tool detects both continuous and non-continuous line changes within one scope.

@@

\paragraph{Running example: path extraction.} Figure~\ref{fig:extraction} illustrates path extraction on a simple change, starting from the ASTDiff. The change (shown in the top right corner of the Figure) replaces \verb+tail+ with \verb+&tail+ and \verb+head+ with \verb+head.clone()+. After the change, the second argument passes a borrow of \verb+tail+ rather than sending the ownership of \verb+tail+ to the callee; and, the third argument sends a clone of \verb+head+ to the callee instead of sending \verb+head+ itself. A DFS on this tree yields three different paths, which we show in red, green, and purple. A path represents a sequence of involved non~terminals. It is crucial to record the order of nodes in the path, so that we can record that order in our datapoint embedding.

@@

The most naive approach would be to only report the observed non~terminals within the diff. That yields a fully order-insensitive representation. In such an embedding, for instance, there is no difference between two nested if statements and two if statements beside each other. On the other hand, theoretically, for full order-sensitiveness, we would need to embed all possible combinations of items from $\mathtt{NT} \cup \mathtt{BCE}$ as our dimensions, which yields a combinatorial explosion; this could potentially be reduced somewhat, but is still impractical. While we can implement a specific workaround to distinguish nested ifs (and we have done so), the general point about the large dimensionality of fully order-sensitive embedding still holds.

A reasonable work around would be to pass from the full set of non~terminals to a smaller set of categories (e.g. a category for larger entities like class or function definitions, and a category for smaller entities like statements and expressions), and then to record all possible combinations of these categories (similar to \cite{hanam2016discovering}). That ad-hoc approach reduces the number of dimensions and provides more reasonable and syntactically-correct combinations. However, it would still yield a sparse dataset.

We propose a novel representation. Similar to the naive approach, we define the columns of our dataset as the set of items $\mathtt{NT} \cup \mathtt{BCE}$ collected from Syn. However, to account for order-sensitivity, we introduce two additional steps. Firstly, we record the number of occurrences of items from $\mathtt{NT} \cup \mathtt{BCE}$ within each distinct root-level scope (e.g. structs, functions, impl blocks, etc). This captures the varying occurrence counts resulting from different element orders in the program. Secondly, we incorporate a weighting scheme, designed semi-automatically, to prioritize non~terminals that we consider more salient for Rust bugs.

\paragraph{Running example: computing fixed-sized vectors.} Figure~\ref{fig:essence} depicts the fixed-sized vectors, as described in this subsection, for the change from Figure~\ref{fig:extraction}. All of the columns in this vector belong to scope \verb+ItemFn+---that is, these non~terminals occur within a function. The Figure also depicts the essence of change, to be described immediately below.

@@

    \caption{\label{fig:essence}Multiplying the number of occurrences of items observed in the tree by their respective weights results in the essence of change, visualized with a circle pack figure at bottom right. All non~terminals here occur within scope {\texttt ItemFn}.}
\end{figure}

Our main requirement is that our datapoints summarize the key changes in ASTDiffs---a datapoint should foreground the most important change in a diff. For instance, the purple (right-most) path in Figure~\ref{fig:extraction} could be described as a change in a local variable declaration; a change in a function call; calling a method of one of the function arguments; or calling clone() on one of the function arguments. The last description, in our opinion, is the most useful one, and we designed our weighting scheme to embed this value judgment.
We made two observations about changes. First, we saw that items (non~terminals or BC-related elements) that occur closer to AST leaves tend to be more semantically important. Second, and relatedly, items closer to the root tend to be repeated in all the paths within one ASTDiff (e.g. ItemFn, Local, ExprCall in Figure~\ref{fig:extraction}). We thus want to prioritize items that tend to occur closer to leaves and de-prioritize items that tend to occur closer to roots.

To design our weighting scheme, we collected empirical data about locations of item occurrences in a corpus. Specifically, we mined 20 of the most recent bug related commits of the same projects described in Table~\ref{table:repos}, and ran them through our pipeline. We recorded the total number of occurrences of each item. Then, if $\# i$ is the number of occurrences of $i$, we gave $i$ a weight of $1/\# i$. Because items that occur closer to the root also occur more often, per our observation, this gives higher weights to items that occur infrequently, i.e. closer to the leaves. Items that occur often and have lower weights are, we believe, less important in describing a change.

@@

\paragraph{Running example: essence of change.} Figure~\ref{fig:essence} shows a complete example for computing the essence of change. The essence of change vector essentially serves as our embedding vector. The Figure shows the process for transforming the patch illustrated in Figure~\ref{fig:extraction} from fixed-size datapoints (computed as described in the previous subsection) into the essence of change vector. To better illustrate the essence vector, Figure~\ref{fig:essence} also graphically presents it using a circle pack figure. The radius of each circle corresponds to the importance of the item inside it.

@@

\subsubsection{\label{sec:manual_analysis_parameter_tuning}Manual analysis for parameter tuning}

@@

\subsubsection{\label{sec:manual_analysis_cluster_selection}Manual analysis for cluster selection}
% 3 types, refactoring dropped
Like \cite{yang2022mining}, and following the definitions of \cite{cotroneo2019analyzing}, we divide the clusters in three different groups:

@@

\begin{lstlisting}[language=Rust, style=colouredRust]
// First Pattern (more important): change clone to borrow
-- foo1(arg.clone);
++ foo1(&arg);

// Second Pattern (less important): change a statement in an fn
fn foo2() {
--  stmt1;
++  stmt2;
}

### 2023-06-14 related.tex
last read through

Researchers have proposed different methods for code repair, each of which can be said to follow a different mindset. Following the taxonomy propsed by \cite{liu2018survey}, we briefly present two mindsets, and will discuss how having a set of predefined fix patterns can aid repair.

@@

Overfitting poses a substantial challenge for automated program repair (APR) systems, where synthesized patches may pass test cases but fail to be correct. Automated patch correctness assessment (APCA) aims to inspect the correctness of generated patches using dynamic and static analyses. While our project has a different objective than APCA, we share similarities in design decisions with APCA systems, particularly in feature extraction. \cite{ye2021automated} propose ODS, an overfitting patch detection system, which assumes that code features capturing universal correctness properties can be utilized to classify overfitting patches across program repair systems and software projects. ODS extracts 202 manually crafted code features by comparing fixed and buggy code and employs supervised learning to build a distributional model for classifying unseen patches, achieving over 70\% accuracy in evaluations on Defects4J, Bugs.jar, and Bears benchmarks. Similarly, our feature extraction involves manual crafting of features. In another study, \cite{tian2022change} present an enhanced APCA model that transforms the problem into a question-answering scenario and incorporates a neural architecture to learn semantic correlations between bug reports and commit messages. Additionally, \cite{lin2022context} introduce \textsc{Cache}, a neural-based context-aware AST embedding method that captures both the change part and the unchanged part (context code) of patches. Evaluation results demonstrate that context knowledge significantly enhances the performance and accuracy of APCA systems.

@@

\cite{hoang2020cc2vec} introduced CC2Vec, a neural network model that utilizes log messages accompanying code changes to learn representations capturing the semantic information. By addressing challenges in distinguishing between added and removed code and incorporating an attention mechanism, CC2Vec outperforms previous techniques for code change representation. The model employs a hierarchical attention network, which incorporates a multi-level bidirectional GRU recurrent neural network, to encode information about changed code tokens, lines, and hunks into vectors representing added and removed changes. These vectors are then combined and fed into a hidden layer for predicting the target function. Evaluation across log message generation, bug fixing patch identification, and just-in-time defect prediction tasks demonstrates the superior performance of CC2Vec compared to state-of-the-art approaches in all three tasks.

Our work tends to be more similar to that of~\cite{hanam2016discovering}. However, our novel program embedding approach differs from what they did on certain key decisions. In~\cite{hanam2016discovering}, they introduced the Feature Properties table, which is a categorization for program elements. Embedding programs based on this table resulted in an order-sensitive embedding, however, it made the datapoints sparse. Our datapoints are more dense than theirs. More specifically, the number of our dimensions is more than 11x smaller.

\subsection{Refactoring tools for Rust}
Clippy is a widely known as the best code refactoring tool for Rust. It can identify over 600 common mistakes and offer automated fixes. It provides users with the flexibility to adjust the linting level, allowing them to control the number of lint suggestions they receive. As discussed in Section~\ref{sec:bc_patterns}, we found that Clippy can detect and address five of our borrow-checker-related patterns. \cite{sam2017automated} introduced a refactoring tool that uses the Rust compiler infrastructure to perform various refactoring operations, including renaming variables, arguments, fields, functions, methods, structs, and enumerations, as well as inlining local variables, and reifying and eliding lifetime parameters. In a more advanced use case, \cite{ling2022rust} proposed CRustS, a tool designed for refactoring C2Rust output. C2Rust is a C to Rust transpiler that converts C code into Rust syntax but retains the unsafe semantics of C. By relaxing the constraint of preserving semantics during transformations, CRustS increases the proportion of transformed code that successfully passes the safety checks of the Rust compiler.

@@

To our knowledge, our work is the first to use an automated mining and code analysis pipeline to find pervasive patterns in Rust open source projects, even if previous studies have investigated common bug patterns in Rust.

\cite{qin2020understanding} conducted the first empirical study on real-world Rust program behaviours. They manually inspected 850 unsafe code usages and 17 bugs across five open-source Rust projects, five Rust libraries, two online security databases, and the Rust standard library. They analyzed the motivation behind unsafe code usage and removal, in addition to recording 70 memory-safety issues and 100 concurrency bugs. They also provided Rust programmers with some suggestions and insights to develop better Rust programs. Using the results of their manual analysis, they designed two bug detectors, and provided recommendations for developing more bug detectors in the future.

\cite{li2021mirchecker} present MirChecker, a fully automated bug detection framework for Rust. This framework works by carrying out static analysis on Rust's Mid-level Intermediate Representation (MIR). The tool exploits the insights obtained from observing existing bugs (by studying reported CVEs) in Rust code bases. Using both numerical (e.g. integer bounds) and symbolic (e.g. modelling memory) information, the framework detects errors by using constraint solving techniques. MirChecker detected 33 new bugs, including 16 memory safety faults, across 12 Rust crates.

>>>>

Researchers have proposed different methods for code repair, each of which can be said to follow a different mindset. Following the taxonomy proposed by \cite{liu2018survey}, we briefly present two mindsets, and will discuss how having a set of predefined fix patterns can aid repair.

@@

Overfitting poses a substantial challenge for automated program repair (APR) systems, where synthesized patches may pass test cases but fail to be correct. Automated patch correctness assessment (APCA) aims to inspect the correctness of generated patches using dynamic and static analyses. While our project has a different objective than APCA, we share similarities in design decisions with APCA systems, particularly in feature extraction. \cite{ye2021automated} propose ODS, an overfitting patch detection system, which assumes that code features capturing universal correctness properties can be utilized to classify overfitting patches across program repair systems and software projects. ODS extracts 202 manually crafted code features by comparing fixed and buggy code, and employs supervised learning to build a distributional model for classifying unseen patches, achieving over 70\% accuracy in evaluations on Defects4J, Bugs.jar, and Bears benchmarks. Similarly, our feature extraction involves manual crafting of features. In another study, \cite{tian2022change} present an enhanced APCA model that transforms the problem into a question-answering scenario and incorporates a neural architecture to learn semantic correlations between bug reports and commit messages. Additionally, \cite{lin2022context} introduce \textsc{Cache}, a neural-based context-aware AST embedding method that captures both the changed part and the unchanged part (context code) of patches. Evaluation results demonstrate that context knowledge significantly enhances the performance and accuracy of systems that attempt to automatically assess patch correctness (which is not our goal).

@@

\cite{hoang2020cc2vec} introduced CC2Vec, a neural network model that utilizes log messages accompanying code changes to learn representations capturing semantic information. By addressing challenges in distinguishing between added and removed code and incorporating an attention mechanism, CC2Vec outperforms previous techniques for code change representation. The model employs a hierarchical attention network, which incorporates a multi-level bidirectional GRU recurrent neural network, to encode information about changed code tokens, lines, and hunks into vectors representing added and removed changes. These vectors are then combined and fed into a hidden layer for predicting the target function. Evaluation across log message generation, bug fixing patch identification, and just-in-time defect prediction tasks demonstrates the superior performance of CC2Vec compared to state-of-the-art approaches in all three tasks.

Our work tends to be more similar to that of~\cite{hanam2016discovering}. However, our novel program embedding approach differs from what they did on certain key decisions. In~\cite{hanam2016discovering}, they introduced the Feature Properties table, which is a categorization for program elements. Embedding programs based on this table resulted in an order-sensitive embedding. However, it made the datapoints sparse, and our datapoints are more dense than theirs. More specifically, the number of our dimensions is more than 11$\times$ smaller.

\subsection{Refactoring tools for Rust}
Clippy is widely known as the best code refactoring tool for Rust. It can identify over 600 common mistakes and offer automated fixes. It provides users with the flexibility to adjust the linting level, allowing them to control the number of lint suggestions they receive. As discussed in Section~\ref{sec:bc_patterns}, we found that Clippy can detect and address five of our borrow-checker-related patterns. \cite{sam2017automated} introduced a refactoring tool that uses the Rust compiler infrastructure to perform various refactoring operations, including renaming variables, arguments, fields, functions, methods, structs, and enumerations, as well as inlining local variables, and reifying and eliding lifetime parameters. In a more advanced use case, \cite{ling2022rust} proposed CRustS, a tool designed for refactoring C2Rust output. C2Rust is a C to Rust transpiler that converts C code into Rust syntax but retains the unsafe semantics of C. By relaxing the constraint of preserving semantics during transformations, CRustS increases the proportion of transformed code that successfully passes the safety checks of the Rust compiler.

@@

To our knowledge, our work is the first to use an automated mining and code analysis pipeline to find pervasive patterns in Rust open source projects, even if previous studies have (manually) investigated common bug patterns in Rust.

\cite{qin2020understanding} conducted the first empirical study on real-world Rust program behaviours. They manually inspected 850 unsafe code usages and 17 bugs across five open-source Rust projects, five Rust libraries, two online security databases, and the Rust standard library. They analyzed the motivation behind unsafe code usage and removal, in addition to recording 70 memory-safety issues and 100 concurrency bugs. They also provided Rust programmers with some suggestions and insights to develop better Rust programs. Using the results of their manual analysis, they designed two bug detectors, and provided recommendations for developing more bug detectors in the future. Our automated approach to finding bug patterns has similar implications to their work but can operate at a far larger scale.

\cite{li2021mirchecker} present MirChecker, a fully automated bug detection framework for Rust. This framework works by carrying out static analysis on Rust's Mid-level Intermediate Representation (MIR). The tool exploits the insights obtained from manually observing existing bugs (by studying reported CVEs) in Rust code bases. Using both numerical (e.g. integer bounds) and symbolic (e.g. modelling memory) information, the framework detects errors by using constraint solving techniques. MirChecker detected 33 new bugs, including 16 memory safety faults, across 12 Rust crates.

### 2023-06-14 results.tex
last read through

Figure~\ref{fig:rq1} shows the final results. Although we found that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 88\% of the time---we investigated the other 12\% as well. We found that our embedding tended to fail in change patterns where the change involves many elements (e.g. adding a whole new \verb+impl+ block and implementing new functions in it). Such failing change patterns tend to apply across projects less often, as having many changed elements make it easier for datapoints to be dissimlar.

@@

\subsubsection{Attributes}
The most common bug fix pattern in our corpus modifies attributes. In Rust, attributes enable developers to add metadata to program elements. Some of the attributes---presumably the vast majority of those involved in bug fixes---affect the semantics of the program; for instance, libraries generate different code based on the attributes that the developer specifies.

Developers can add attributes to structs, struct fields, enum variants, match expression arms, etc.
%I don't think we're well served to talk about inner vs outer attributes, not relevant to readers
%There are two types of attributes\todo{how important is inner versus outer?}: Inner Attributes and Outer Attributes. Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute.
The concept of attributes in Rust is similar to Python's decorators and Java's annotations. While each language may have its own syntax and specific use cases, the fundamental idea of attaching metadata to code elements is shared across these languages. Some attributes, decorators, and annotations have no run-time effect and thus could not cause bugs (if we accept a purely operational definition of ``bug''). \cite{yang2022mining}~reported no bug patterns involving Python's decorators in their Python dataset. As we alluded to above, however, Rust has many attributes that definitely have run-time effects. We will see some of these attributes in the following examples.

>>>>

Figure~\ref{fig:rq1} shows the final results. Although we found that our results were acceptable---the result reported by our embedding was top-4 as ranked by humans 88\% of the time---we investigated the other 12\% as well. We found that our embedding tended to fail in change patterns where the change involves many elements (e.g. adding a whole new \verb+impl+ block and implementing new functions in it). Such failing change patterns tend to apply across projects less often, as having many changed elements make it easier for datapoints to be dissimilar.

@@

\subsubsection{Attributes}
A quite common bug fix pattern in our corpus modifies attributes. In Rust, attributes enable developers to add metadata to program elements. Some of the attributes---presumably the vast majority of those involved in bug fixes---affect the semantics of the program; for instance, libraries generate different code based on the attributes that the developer specifies.

Developers can add attributes to structs, struct fields, enum variants, match expression arms, etc.
%I don't think we're well served to talk about inner vs outer attributes, not relevant to readers
%There are two types of attributes\todo{how important is inner versus outer?}: Inner Attributes and Outer Attributes. Inner Attributes apply to the program element encapsulating them, while Outer Attributes apply to the program element coming after the attribute.
The concept of attributes in Rust is similar to Python's decorators and Java's annotations. While each language may have its own syntax and specific use cases, the fundamental idea of attaching metadata to code elements is shared across these languages. Some attributes, decorators, and annotations have no run-time effect and thus could not cause bugs (if we accept a purely operational definition of ``bug''). \cite{yang2022mining}~reported no bug patterns involving Python's decorators in their Python dataset. As we alluded to above, however, Rust has many attributes that definitely have run-time effects. We will see some of these attributes in the following examples. \\
